A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2013-10-09 jacobian_calc_CMAME.tex
removed some speculation

\subsection{Thoughts on Results} Based on the experimental results, it appears that for users of
\emph{Peridigm} there is a case to be made for using automatic-differentiation for reasons of
performance and accuracy. Although, the greater speed of finite-difference for certain multicore
tests done in this study may give some users pause. While the results showed that complex-step
produced more accurate tangent-stiffness matrices than central-difference and forward-difference 
under the parameters of the tests, it was not determined whether or not this is a clear advantage of complex-step over
finite-difference in terms of accuracy of final predicted node positions and speed of convergence. For other
simulation software it is obviously necessary to collect the same types measurements as decribed in
this paper in order to make a design decision. For new developments it is recommended that combined
accuracy and speed of iteration mesurments be conduted along with entirely separate convergence
studies. For reasons explained in the next subsection, a convergence study in \emph{Peridigm} performed using
the implementations of the methods used for this paper would be moot.

\subsection{Thoughts on \emph{Perdigm}}

After the bulk of the study had been conducted, gained familiarity with \emph{Peridigm} revealed
that the implementations of forward-difference, central-difference and by virtue of reusing code,
complex-step, in \emph{Peridigm} were not optimal due to inefficiencies in the implementations of
the Linear Peridynamic Solid (LPS) model. Each of these implementations included a large amount of
unnecessary re-calculation of constant values.  The approach used does minimize temporary memory
use, and allow other objects in \emph{Peridigm} to make calls directly to methods in the material
objects, but is wastefull. Unpublished preliminary results show that when calculations for the
constant quantity 'zeta' in the LPS implementation are re-used and only perturbed and previously
pertubed functions of independent variables are recomputed throughout the course of a Jacobian
evaluation, complex-step performs at the speed of unmodified forward-difference. Because of this it
is speculated that forward-difference and central-difference could be modified in nearly the same
way to perform more quickly than AD. In fact, because central-difference involves double the number
of model evaluations than that of complex-step, central-difference stands to gain proportionally
more efficiency. Although no formal academic paper will be likely written based on this, a brief
report and an enhanced performance implementation of finite-difference for peridynamic LPS Jacbian
calculation  in \emph{Peridigm} will eventually be made available on the corresponding author's
website and is intended to be submitted for official inclusion in future releases of \emph{Peridigm}.

>>>>

\subsection{Thoughts on Results} Based on the experimental results, it appears that for users of
\emph{Peridigm} there is a case to be made for using automatic-differentiation for reasons of
performance and accuracy. Although, the greater speed of finite-difference for certain multicore
tests done in this study may give some users pause. While the results showed that complex-step
produced more accurate tangent-stiffness matrices than central-difference and forward-difference 
under the parameters of the tests, it was not determined whether or not this is a clear advantage of complex-step over
finite-difference in terms of accuracy of final predicted node positions and speed of convergence. For other
simulation software it is obviously necessary to collect the same types measurements as decribed in
this paper in order to make a design decision. For new developments it is recommended that combined
accuracy and speed of iteration mesurments be conduted along with entirely separate convergence
studies.
